Analysing the Performance of Shallow-Water Models with the Roofline Model

### 2016-02-03 docs/papers/roofline_model/gocean_roofline.tex
roofline: change to grace plot of perf and bw of daxpy kernels. Add non-aligned daxpy result to daxpy roofline.
roofline: add daxpy perf. for main-memory problem size. Add discussion of daxpy roofline plot

\begin{figure}
\includegraphics[width=120mm]{daxpy_plot}
\caption{Bandwidth obtained on the E5-1620 CPU by the original and
  modified versions of the daxpy kernel from the STREAM2 benchmark.}
\label{FIG_daxpy_bw}
\end{figure}

@@

Peak absolute performance is obtained for the 256 problem size with
SSE instructions. The 512 problem size spills from L3 cache and
becomes bound by the bandwidth to main memory: SIMD vectorisation
gives no performance benefit. The fact that best performance is
obtained when the problem size fits within L3 indicates that there are
other overheads that dominate when the problem size is small enough to
fit within L2. This is possibly because the inner, SIMD-vectorised
loop then only has a trip-count of 35. (This is not 32 because of the way
the simulation domain is set-up; it is the {\em internal} region of
the domain that has dimension $32^2$.) {\bf Remainder and xxx loops.}

@@

As well as examining the performance of the Continuity kernel, we have
also experimented with altering the DAXPY kernel of the STREAM2
benchmark so as to vary its $AI$. We did this by adding further
arithmetic operations but making use of the same array
elements. We used four different kernels:
\begin{enumerate}
\item {\tt b(i) = b(i) + scalar*a(i)}
\item {\tt b(i) = b(i) + scalar*a(i)+ a(i)*b(i)}
\item {\tt b(i) = b(i) + scalar*a(i)+ a(i)*b(i)*b(i)}
\item {\tt b(i) = b(i) + scalar*a(i)+ a(i)*b(i)*b(i)*b(i)}
\end{enumerate}
Figure~\ref{FIG_daxpy_roofline} shows the performance of these kernels
on a roofline plot for the E5-1620 CPU. They generally reach L2 and L3
bandwidth apart from at the lowest AI. Continuity kernel has AI of
{\bf xxx more discussion here}

\begin{figure}
\includegraphics[width=120mm]{roofline_daxpy}
\caption{Roofline plot of daxpy-based kernels on the E5-1620 CPU. Each
  cluster of bars represents a different kernel. Bars within a cluster
  are for problem sizes chosen to fit within the L3, L2 and L1 cache
  (arrays of 94,427, 8271 and 1533 elements, respectively). }
\label{FIG_daxpy_roofline}
\end{figure}

>>>>

\begin{figure}
\includegraphics[width=120mm]{daxpy_bw_perf}
\caption{Bandwidth obtained on the E5-1620 CPU by the original and
  modified versions of the daxpy kernel from the STREAM2 benchmark.}
\label{FIG_daxpy_bw}
\end{figure}

@@

Peak absolute performance is obtained for the 256 problem size with
SSE instructions. The 512 problem size spills from L3 cache and
becomes bound by the bandwidth to main memory: SIMD vectorisation
gives no performance benefit. The fact that best performance is
obtained when the problem size fits within L3 indicates that there are
other overheads that dominate when the problem size is small enough to
fit within L2. This is possibly because the inner, SIMD-vectorised
loop then only has a trip-count of 35. (This is not 32 because of the way
the simulation domain is set-up; it is the {\em internal} region of
the domain that has dimension $32^2$.) {\bf Remainder and peel loops.}

@@

As well as examining the performance of the Continuity kernel, we have
also experimented with altering the DAXPY kernel of the STREAM2
benchmark so as to vary its $AI$. We did this by adding further
arithmetic operations but making use of the same array
elements. We used five different kernels:
\begin{enumerate}
\item {\tt y(i) = y(i) + a*x(i)}, AI = 0.125
\item {\tt y(i) = y(i) + a*x(i)+ x(i)*y(i)}, AI = 0.167
\item {\tt y(i) = y(i) + a*x(i)+ x(i)*y(i)*y(i)}, AI = 0.208
\item {\tt y(i) = y(i) + a*x(i)+ x(i)*y(i)*y(i)*y(i)}, AI = 0.25
\item {\tt y(i) = y(i) + a*x(i)+ x(i-1)*y(i)*y(i)*y(i)}, AI = 0.25
\end{enumerate}
The fifth of these introduces an un-aligned array access ({\tt
  x(i-1)}) so as to better represent real-life finite difference
kernels.  Figure~\ref{FIG_daxpy_roofline} shows the performance of
these kernels on a roofline plot for the E5-1620 CPU. The performance
of each kernel is plotted for four problem sizes, chosen so as to fit
within L1, L2, L3 cache and main memory. The performance of kernels 2,
3 and 4 is entirely consistent with the roofline model for problem
sizes that fit within the L3 and L3 caches (red and black bars). Their
performance for the smallest problem size doesn't quite reach the L1
bandwidth limit but does appear to track the roofline.  Kernel 1 has
the lowest AI of the five kernels and this appears to limit its
performance; at no point does the performance reach the limit imposed
by the bandwidth to the relevant level of the memory hierarchy. {\bf need to expore the reason for this}

The performance of kernel 5 is plotted using patterned bars in
Figure~\ref{FIG_daxpy_roofline}. This kernel has the same AI as kernel
4 but has an un-aligned array access. Comparing the performance of
these two kernels we see that the resulting reduction in the
efficiency with which the kernel can be SIMD vectorised (because of
the need for gather/scatter operations) harms performance,
particularly when the cost of accessing memory is small (i.e. when the
problem is contained in L1 or L2). Since all of the NEMOLite2D kernels
contain un-aligned array accesses, this goes some way to accounting
for the fact that they do not reach the theoretical limits of the
roofline plot in Figure~\ref{FIG_roofline_haswell}.

\begin{figure}
\includegraphics[width=120mm]{roofline_daxpy}
\caption{Roofline plot of daxpy-based kernels on the E5-1620 CPU. Each
  cluster of bars represents a different kernel. The green, black, red and
  blue bars are for problem sizes chosen to fit within main memory, L3, L2 and
  L1 caches (arrays of 3.3M, 94K, 8.3K and 1.5K elements),
  respectively. Bars with patterned fill are for kernel 5 (unaligned
  access).}
\label{FIG_daxpy_roofline}
\end{figure}
